---
title: "2020-04-16 - Exploratory Experiments - Analysis"
output:
  html_document:
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
---

## Analysis Dependencies

```{r, message=FALSE}
library(ggplot2)  # (Wickham, 2016)
library(tidyr)    # (Wickham and Henry, 2020)
library(dplyr)    # (Wickham et al., 2020)
library(reshape2) # (Wickham, 2007)
library(cowplot)  # (Wilke, 2019)
library(reticulate) 
```

We conducted these analyses using the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
os <- import("os") # I'm a monster
data_dir <- "./data"
data_path <- os$path$join(data_dir, "agg_data.csv")
agg_data <- read.csv(data_path, na.strings="NONE")

agg_data$BIT_FLIP_PROB <- as.factor(agg_data$BIT_FLIP_PROB)
agg_data$DRIFT <- agg_data$TOURNAMENT_SIZE==1

# Label
chg_rate_label <- function(mag, interval, drift) {
  if (drift) { return("drift") }
  else if (interval == 0) { return("0") }
  else { return(paste(mag, interval, sep="/")) }
}

agg_data$chg_rate_label <- factor(mapply(chg_rate_label, 
                                            agg_data$CHANGE_MAGNITUDE,
                                            agg_data$CHANGE_FREQUENCY,
                                            agg_data$DRIFT),
                                     levels=c("drift", "0", "1/256", "1/128",
                                              "1/64", "1/32", "1/16", "1/8",
                                              "1/4", "1/2", "1/1", "2/1", 
                                              "4/1", "8/1", "16/1", "32/1",
                                              "64/1", "128/1", "256/1", 
                                              "512/1", "1024/1", "2048/1",
                                              "4096/1"))

theme_set(theme_cowplot())

# Divide up the data into convenient partitions
data_gradient_phase0 <- 
  filter(agg_data,
         GRADIENT_MODEL==1 & evo_phase == 0 & update == 50000)
data_gradient_phase1 <-
  filter(agg_data,
         GRADIENT_MODEL==1 & evo_phase == 1 & update == 60000)
data_nk_phase0 <-
  filter(agg_data,
         GRADIENT_MODEL==0 & evo_phase == 0 & update == 50000)
data_nk_phase1 <-
  filter(agg_data,
         GRADIENT_MODEL==0 & evo_phase == 1 & update == 60000)
```

## Evolution Experiment

Evolved genomes for 50,000 generations.

### Coding Sites by Mutation Rate and Environmental Change Rate

#### Gradient fitness model

Facets on mutation rate.

```{r}
ggplot(data_gradient_phase0,
      aes(x=chg_rate_label, y=coding_sites, color=chg_rate_label)) +
  geom_boxplot() +
  facet_wrap(~ BIT_FLIP_PROB) +
  xlab("Environmental change rate") +
  scale_y_continuous(name="# coding sites in best organism",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 10)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=90))
```

Facets on environmental change rate.

```{r}
ggplot(data_gradient_phase0,
      aes(x=BIT_FLIP_PROB, y=coding_sites, color=BIT_FLIP_PROB)) +
  geom_boxplot() +
  facet_wrap(~ chg_rate_label, ncol=3) +
  xlab("Bit substitution rate") +
  scale_y_continuous(name="# coding sites in best organism",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 10)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=90))
```

#### NK fitness model

Facets on mutation rate.

```{r}
ggplot(data_nk_phase0,
      aes(x=chg_rate_label, y=coding_sites, color=chg_rate_label)) +
  geom_boxplot() +
  facet_wrap(~ BIT_FLIP_PROB) +
  xlab("Environmental change rate") +
  scale_y_continuous(name="# coding sites in best organism",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 10)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=90))
```

Facets on environmental change rate.

```{r}
ggplot(data_nk_phase0,
      aes(x=BIT_FLIP_PROB, y=coding_sites, color=BIT_FLIP_PROB)) +
  geom_boxplot() +
  facet_wrap(~ chg_rate_label, ncol=3) +
  xlab("Bit substitution rate") +
  scale_y_continuous(name="# coding sites in best organism",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 10)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=90))
```

### Fitness

Note that in a changing environment, lineages have not necessarily had time to adapt to the environment they are measured against.

#### Gradient fitness model

Faceted by mutation rate.

```{r}
ggplot(data_gradient_phase0,
      aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  facet_wrap(~ BIT_FLIP_PROB) +
  xlab("Environment bits flipped per generation") +
  theme(legend.position="none",
    axis.text.x = element_text(angle=90))
```

Faceted by change rate.

```{r}
ggplot(data_gradient_phase0,
      aes(x=BIT_FLIP_PROB, y=fitness, color=BIT_FLIP_PROB)) +
  geom_boxplot() +
  facet_wrap(~ chg_rate_label) +
  xlab("Bit substitution rate") +
  theme(legend.position="none",
    axis.text.x = element_text(angle=90))
```

#### NK fitness model

Faceted by mutation rate.

```{r}
ggplot(data_nk_phase0,
      aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  facet_wrap(~ BIT_FLIP_PROB) +
  xlab("Environment bits flipped per generation") +
  theme(legend.position="none",
    axis.text.x = element_text(angle=90))
```

Faceted by change rate.

```{r}
ggplot(data_nk_phase0,
      aes(x=BIT_FLIP_PROB, y=fitness, color=BIT_FLIP_PROB)) +
  geom_boxplot() +
  facet_wrap(~ chg_rate_label) +
  xlab("Bit substitution rate") +
  theme(legend.position="none",
    axis.text.x = element_text(angle=90))
```

## Transfer Experiment

We transfer each population into a new (random) static environment to evolve for 10,000 generations. We disallow mutations that alter a genome's genetic architecture. 

### Fitness

#### Gradient fitness model

```{r}
ggplot(data_gradient_phase1,
      aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  facet_wrap(~ BIT_FLIP_PROB) +
  theme(legend.position="none")
```

#### NK fitness model

```{r}
ggplot(data_nk_phase1,
      aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  facet_wrap(~ BIT_FLIP_PROB) +
  theme(legend.position="none")
```

### Coding Sites by Fitness

#### Gradient fitness model

```{r}
ggplot(data_gradient_phase1,
      aes(x=coding_sites, y=fitness, color=chg_rate_label)) +
  geom_point(alpha=0.1) +
  facet_wrap(~ BIT_FLIP_PROB) +
  theme(legend.position="bottom")
```

#### NK fitness model

```{r}
ggplot(data_nk_phase1,
      aes(x=coding_sites, y=fitness, color=chg_rate_label)) +
  geom_point(alpha=0.1) +
  facet_wrap(~ BIT_FLIP_PROB) +
  theme(legend.position="bottom")
```