---
title: "2020-05-22 - Lineage Summary Analyses"
output:
  html_document:
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
---

## Analysis Dependencies

```{r, message=FALSE}
library(ggplot2)  # (Wickham, 2016)
library(tidyr)    # (Wickham and Henry, 2020)
library(dplyr)    # (Wickham et al., 2020)
library(reshape2) # (Wickham, 2007)
library(cowplot)  # (Wilke, 2019)
library(patchwork)
```

We conducted these analyses using the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
theme_set(theme_cowplot())
data_path <- "./data-summary/lineages_summary.csv"
agg_data <- read.csv(data_path, na.strings="NONE")

agg_data$BIT_FLIP_PROB <- as.factor(agg_data$BIT_FLIP_PROB)
agg_data$DRIFT <- agg_data$TOURNAMENT_SIZE==1

# Label
chg_rate_label <- function(mag, interval, drift) {
  if (drift) { return("drift") }
  else if (interval == 0) { return("0") }
  else { return(paste(mag, interval, sep="/")) }
}

agg_data$chg_rate_label <- factor(mapply(chg_rate_label, 
                                            agg_data$CHANGE_MAGNITUDE,
                                            agg_data$CHANGE_FREQUENCY,
                                            agg_data$DRIFT),
                                     levels=c("drift", "0", "1/256", "1/128",
                                              "1/64", "1/32", "1/16", "1/8",
                                              "1/4", "1/2", "1/1", "2/1", 
                                              "4/1", "8/1", "16/1", "32/1",
                                              "64/1", "128/1", "256/1", 
                                              "512/1", "1024/1", "2048/1",
                                              "4096/1"))

agg_data$mutations_per_step <- agg_data$accum_muts / agg_data$length
gradient_data <- filter(agg_data, GRADIENT_MODEL==1)
nk_data <- filter(agg_data, GRADIENT_MODEL==0)
```

## Analyses 

### Lineage Length

```{r}
g <- 
ggplot(gradient_data, aes(x=chg_rate_label, y=length, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylab("Lineage Length") +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position = "none") + 
  ggsave("imgs/lineage-length_gradient.pdf", width=15, height=10)

n <- 
ggplot(nk_data, aes(x=chg_rate_label, y=length, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylab("Lineage Length") +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("NK Fitness Model") +
  theme(legend.position = "none") + 
  ggsave("imgs/lineage-length_nk.pdf", width=15, height=10)

(g / n) + ggsave("imgs/lineage-length.pdf", width=15, height = 15)
```

### Total Accumulated Mutations

```{r}
g <-
ggplot(gradient_data, aes(x=chg_rate_label, y=accum_muts, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylab("Accumulated\nMutations") +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position = "none") +
  ggsave("imgs/total-mutation-accumulation_gradient.pdf", width=15, height=10)

n <-
ggplot(nk_data, aes(x=chg_rate_label, y=accum_muts, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylab("Accumulated\nMutations") +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("NK Fitness Model") +
  theme(legend.position = "none") +
  ggsave("imgs/total-mutation-accumulation_nk.pdf", width=15, height=10)

(g / n) + ggsave("imgs/total-mutation-accumulation.pdf", width=15, height = 15)
```

### Avergage mutations accumulated per lineage-step

Here, we normalize mutations accumulated by lineage length.

```{r}
g <- 
ggplot(gradient_data, 
       aes(x=chg_rate_label, y=mutations_per_step, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylab("Accumulated Mutations / Step") +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position = "none") 

n <- 
ggplot(nk_data, 
       aes(x=chg_rate_label, y=mutations_per_step, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylab("Accumulated Mutations / Step") +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("NK Fitness Model") +
  theme(legend.position = "none")

(g/n) + ggsave("imgs/total-accumulation-per-step.pdf", width = 15, height = 15)
```

### Accumulated Mutation Distribution

Melt accumulations by type to make plotting easier.

```{r}
gradient_melted <- melt(gradient_data,
                        measure.vars=c("accum_gene_move_muts",
                                       "accum_bit_flip_muts",
                                       "accum_bit_ins_muts",
                                       "accum_bit_del_muts"),
                        variable.name="mutation_type",
                        value.name="mutation_type_accum")
gradient_melted$mutation_type_accum_per_step <- 
  gradient_melted$mutation_type_accum / gradient_melted$length
gradient_melted$mutation_type_proportion <-
  gradient_melted$mutation_type_accum / gradient_melted$accum_muts

nk_melted <- melt(nk_data,
                  measure.vars=c("accum_gene_move_muts",
                                 "accum_bit_flip_muts",
                                 "accum_bit_ins_muts",
                                 "accum_bit_del_muts"),
                  variable.name="mutation_type",
                  value.name="mutation_type_accum")
nk_melted$mutation_type_accum_per_step <- 
  nk_melted$mutation_type_accum / nk_melted$length
nk_melted$mutation_type_proportion <- 
  nk_melted$mutation_type_accum / nk_melted$accum_muts
```

#### Mutation Type Proportions

```{r}
g <- 
ggplot(gradient_melted, aes(x=mutation_type, y=mutation_type_proportion, color=mutation_type)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylim(0.0, 1.0) +
  facet_grid(BIT_FLIP_PROB~chg_rate_label) +
  ggtitle("Gradient Fitness Model") +
  scale_x_discrete(name="Mutation Type",
                 breaks=c("accum_gene_move_muts",
                          "accum_bit_flip_muts",
                          "accum_bit_ins_muts",
                          "accum_bit_del_muts"),
                 label=c("Gene Move",
                         "Bit Flip",
                         "Bit Insertion",
                         "Bit Deletion")) +
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill=NA)
        ) + 
  ggsave("imgs/mutation-type-proportions_gradient.pdf", width=20, height=10)

n <- 
ggplot(nk_melted, aes(x=mutation_type, y=mutation_type_proportion, color=mutation_type)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  ylim(0.0, 1.0) +
  scale_x_discrete(name="Mutation Type",
                   breaks=c("accum_gene_move_muts",
                            "accum_bit_flip_muts",
                            "accum_bit_ins_muts",
                            "accum_bit_del_muts"),
                   label=c("Gene Move",
                           "Bit Flip",
                           "Bit Insertion",
                           "Bit Deletion")) +
  facet_grid(BIT_FLIP_PROB~chg_rate_label) +
  ggtitle("NK Fitness Model") +
  theme(legend.position = "none",
        panel.border = element_rect(colour = "black", fill=NA)
        ) + 
  ggsave("imgs/mutation-type-proportions_nk.pdf", width=20, height=10)

(g/n) + ggsave("imgs/mutation-type-proportions.pdf", width=20, height=20)

```

As a stacked bar chart:

```{r}
# Stacked bar charts?
summary <- read.csv(text="cfg_id,BIT_FLIP_PROB,chg_rate_label,GRADIENT_MODEL,accum_muts,accum_moves,accum_flips,accum_ins,accum_dels")
mutation_rates = c(0.003, 0.03)
change_rates = c("drift", "0", "1/4", "64/1")
fitness_model = c(0, 1)
combo_id <- 0
for (g_model in fitness_model) {
  for (mut_rate in mutation_rates) {
    for (chg_rate in change_rates) {
      temp <- filter(agg_data,
                     GRADIENT_MODEL==g_model 
                     & BIT_FLIP_PROB==mut_rate
                     & chg_rate_label==chg_rate)
      if (length(temp$SEED) == 0) { next(); } # Skip cross sections w/no data
      total <- sum(temp$accum_muts)
      moves <- sum(temp$accum_gene_move_muts)
      flips <- sum(temp$accum_bit_flip_muts)
      ins <- sum(temp$accum_bit_ins_muts)
      dels <- sum(temp$accum_bit_del_muts)
      df <- data.frame(cfg_id=c(combo_id),
                 BIT_FLIP_PROB=c(mut_rate),
                 chg_rate_label=c(chg_rate),
                 GRADIENT_MODEL=c(g_model),
                 accum_muts=c(total),
                 accum_moves=c(moves),
                 accum_flips=c(flips),
                 accum_ins=c(ins),
                 accum_dels=c(dels))
      summary <- rbind(summary, df)
      combo_id <- combo_id + 1
    }
  }
}
melted_summary <- melt(summary,                  
                       measure.vars=c("accum_moves",
                                      "accum_flips",
                                      "accum_ins",
                                      "accum_dels"),
                      variable.name="mutation_type",
                      value.name="mutation_type_accum")
melted_summary$mut_type_proportion <- melted_summary$mutation_type_accum / melted_summary$accum_muts


g <-
ggplot(filter(melted_summary, GRADIENT_MODEL==1), 
       aes(fill=mutation_type, y=mut_type_proportion, x=chg_rate_label)) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_discrete(name="Mutation Type",
                      breaks=c("accum_moves",
                            "accum_flips",
                            "accum_ins",
                            "accum_dels"),
                      label=c("Gene Move",
                              "Bit Flip",
                              "Bit Insertion",
                              "Bit Deletion")) +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position = "right")

n <- 
ggplot(filter(melted_summary, GRADIENT_MODEL==0), 
       aes(fill=mutation_type, y=mut_type_proportion, x=chg_rate_label)) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_discrete(name="Mutation Type",
                      breaks=c("accum_moves",
                            "accum_flips",
                            "accum_ins",
                            "accum_dels"),
                      label=c("Gene Move",
                              "Bit Flip",
                              "Bit Insertion",
                              "Bit Deletion")) +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("NK Fitness Model") +
  theme(legend.position = "right") 

(g / n) + ggsave("imgs/summary-mutation-type-proportions.pdf", width=15, height=10)
```

What are the relative importances of different mutation types for each environmental? I.e., are there a higher proportion of genetic architecture altering mutations in changing environments versus static environments?

```{r}
# Low mutation rate
mut_rates <- c("0.003", "0.03")
mut_types <- c("accum_gene_move_muts", 
               "accum_bit_flip_muts",
               "accum_bit_ins_muts",
               "accum_bit_del_muts")
for (mut_type in mut_types) {
  for (mut_rate in mut_rates) {
    print("==========================================================")
    print(paste("Mutation type:", mut_type, "Mutation rate:", mut_rate))
    df <- filter(gradient_melted, 
                 BIT_FLIP_PROB==mut_rate & mutation_type==mut_type)
    
    print(paste(
      "Median of drift",
      median(filter(df, chg_rate_label=="drift")$mutation_type_proportion)
    ))
    print(paste(
      "Median of static",
      median(filter(df, chg_rate_label=="0")$mutation_type_proportion)
    ))
    print(paste(
      "Median of 1/4",
      median(filter(df, chg_rate_label=="1/4")$mutation_type_proportion)
    ))
    
    kt <- 
    kruskal.test(
      formula = mutation_type_proportion ~ chg_rate_label,
      data = df
    )
    print(kt)
    if (kt$p.value < 0.05) {
      wt <-
      pairwise.wilcox.test(
        x=df$mutation_type_proportion,
        g=df$chg_rate_label,
        p.adjust.method="bonferroni",
        conf.int=TRUE
      )
      print(wt)
    }
  }
}
```

#### Gene movement mutations accumulated per-step

```{r}
g <- 
ggplot(filter(gradient_melted, mutation_type=="accum_gene_move_muts" & chg_rate_label!="drift"), aes(x=chg_rate_label, y=mutation_type_accum_per_step, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("Gradient Fitness Model - Gene Move Accumulation (per step)") +
  theme(legend.position = "none")

n <- 
ggplot(filter(nk_melted, mutation_type=="accum_gene_move_muts" & chg_rate_label!="drift"), aes(x=chg_rate_label, y=mutation_type_accum_per_step, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("NK Fitness Model - Gene Move Accumulation (per step)") +
  theme(legend.position = "none")

(g/n) + ggsave("imgs/gene-move-accumulation-per-step.pdf", width=20, height=10)
```

#### Gene move mutation proportions

```{r}
g <- 
ggplot(filter(gradient_melted, mutation_type=="accum_gene_move_muts" & chg_rate_label!="drift"), aes(x=chg_rate_label, y=mutation_type_proportion, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("Gradient Fitness Model - Gene Move Proportion") +
  theme(legend.position = "none")

n <- 
ggplot(filter(nk_melted, mutation_type=="accum_gene_move_muts" & chg_rate_label!="drift"), aes(x=chg_rate_label, y=mutation_type_proportion, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  facet_wrap(~BIT_FLIP_PROB) +
  ggtitle("NK Fitness Model - Gene Move Proportion") +
  theme(legend.position = "none")

(g/n) + ggsave("imgs/gene-move-proportion.pdf", width=20, height=10)


```

