---
title: "2020-07-04 - Environment Change Rate Sweeps - Experiment Analysis"
output:
  html_document:
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
---

## Analysis Dependencies

```{r, message=FALSE}
library(ggplot2)  # (Wickham, 2016)
library(tidyr)    # (Wickham and Henry, 2020)
library(dplyr)    # (Wickham et al., 2020)
library(reshape2) # (Wickham, 2007)
library(cowplot)  # (Wilke, 2019)
library(hexbin)   # (Carr, Lewin-Kog, Maechler, and Sarkar, 2020)
```


We conducted these analyses using the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
theme_set(theme_cowplot())
alpha <- 0.05

data_path <- "./data/agg_data.csv"
agg_data <- read.csv(data_path, na.strings="NONE")

agg_data$BIT_FLIP_PROB <- as.factor(agg_data$BIT_FLIP_PROB)
agg_data$DRIFT <- agg_data$TOURNAMENT_SIZE==1

# Label
chg_rate_label <- function(mag, interval, drift) {
  if (drift) { return("drift") }
  else if (interval == 0) { return("0") }
  else { return(paste(mag, interval, sep="/")) }
}

agg_data$chg_rate_label <- factor(mapply(chg_rate_label, 
                                            agg_data$CHANGE_MAGNITUDE,
                                            agg_data$CHANGE_FREQUENCY,
                                            agg_data$DRIFT),
                                     levels=c("drift", "0", "1/256", "1/128",
                                              "1/64", "1/32", "1/16", "1/8",
                                              "1/4", "1/2", "1/1", "2/1", 
                                              "4/1", "8/1", "16/1", "32/1",
                                              "64/1", "128/1", "256/1", 
                                              "512/1", "1024/1", "2048/1",
                                              "4096/1"))

# Divide up the data into convenient partitions
data_gradient_phase0 <- 
  filter(agg_data,
         GRADIENT_MODEL==1 & evo_phase == 0 & update == 50000)
data_gradient_phase1 <-
  filter(agg_data,
         GRADIENT_MODEL==1 & evo_phase == 1 & update == 60000)
data_nk_phase0 <-
  filter(agg_data,
         GRADIENT_MODEL==0 & evo_phase == 0 & update == 50000)
data_nk_phase1 <-
  filter(agg_data,
         GRADIENT_MODEL==0 & evo_phase == 1 & update == 60000)
```

## Evolution Experiment

We evolved populations for 50,000 generations under a range of environmental change rates.

### Fitnesses

Note issue with fitnesses for changing environments.

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase0,
       aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  xlab("Environmental Change") +
  scale_y_continuous(name="Fitness (of best organism)") +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="none")
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase0,
       aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  xlab("Env. bits flipped per generation") +
  scale_y_continuous(name="Fitness  (of best organism)") +
  ggtitle("NK Fitness Model") +
  theme(legend.position="none")
```

### Coding Sites

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase0,
       aes(x=chg_rate_label, y=coding_sites, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(aes(color=chg_rate_label), alpha=0.1) +
  xlab("Environmental Change") +
  scale_y_continuous(name="Coding Sites (in best genotype)",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="none")
```

```{r}
kt <- kruskal.test(formula = coding_sites ~ chg_rate_label, data=data_gradient_phase0)
kt
if (kt$p.value <= alpha) {
  wt <- pairwise.wilcox.test(x=data_gradient_phase0$coding_sites,
                             g=data_gradient_phase0$chg_rate_label,
                             exact=FALSE,
                             p.adjust.method = "bonferroni")
  wt
}
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase0,
       aes(x=chg_rate_label, y=coding_sites, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  scale_y_continuous(name="# coding sites in best organism",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  ggtitle("NK Fitness Model") +
  theme(legend.position="none")
```

```{r}
kt <- kruskal.test(formula = coding_sites ~ chg_rate_label, data=data_nk_phase0)
kt
if (kt$p.value <= alpha) {
  wt <- pairwise.wilcox.test(x=data_nk_phase0$coding_sites,
                             g=data_nk_phase0$chg_rate_label,
                             exact=FALSE,
                             p.adjust.method = "bonferroni")
  wt
}
```

### Genome Length

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase0,
       aes(x=chg_rate_label, y=genome_length, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  ylab("Genome Length") +
  ylim(0, 1025) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="none")
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase0,
       aes(x=chg_rate_label, y=genome_length, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  ylab("Genome Length") +
  ylim(0, 1024) +
  ggtitle("NK Fitness Model") +
  theme(legend.position="none")
```

### Neutral Sites

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase0,
       aes(x=chg_rate_label, y=neutral_sites, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  ylab("Neutral Sites") +
  ylim(-1, 1025) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="none")
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase0,
       aes(x=chg_rate_label, y=neutral_sites, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  ylab("Neutral Sites") +
  ylim(-1, 1025) + # jitter can jitter below 0
  ggtitle("NK Fitness Model") +
  theme(legend.position="none")
```

## Transfer experiment

We lock-down genetic architectures and evolve for 10,000 generations in a random static environment.

### Fitnesses

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase1,
       aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  scale_y_continuous(name="Fitness") +
  ggtitle("Gradient Fitness Model (transfer)") +
  theme(legend.position="none")
```

```{r}
kt <- kruskal.test(formula = fitness ~ chg_rate_label, data=data_gradient_phase1)
kt
if (kt$p.value <= alpha) {
  wt <- pairwise.wilcox.test(x=data_gradient_phase1$fitness,
                             g=data_gradient_phase1$chg_rate_label,
                             exact=FALSE,
                             p.adjust.method = "bonferroni")
  wt
}
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase1,
       aes(x=chg_rate_label, y=fitness, color=chg_rate_label)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  scale_y_continuous(name="Fitness") +
  ggtitle("NK Fitness Model (transfer)") +
  theme(legend.position="none")
```

```{r}
kt <- kruskal.test(formula = fitness ~ chg_rate_label, data=data_nk_phase1)
kt
if (kt$p.value <= alpha) {
  wt <- pairwise.wilcox.test(x=data_nk_phase1$fitness,
                             g=data_nk_phase1$chg_rate_label,
                             exact=FALSE,
                             p.adjust.method = "bonferroni")
  wt
}
```

### Fitness vs Number of Coding Sites

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase1,
       aes(x=coding_sites, y=fitness)) +
  geom_point(aes(color=chg_rate_label)) +
  xlab("Coding Sites") +
  ylab("Fitness") +
  theme(legend.position="top") +
  ggtitle("Gradient Fitness Model")
```

```{r}
cor.test(x=data_gradient_phase1$coding_sites, 
         y=data_gradient_phase1$fitness,
         method="spearman",
         exact=FALSE)
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase1,
       aes(x=coding_sites, y=fitness)) +
  geom_point(aes(color=chg_rate_label)) +
  xlab("Coding Sites") +
  ylab("Fitness") +
  theme(legend.position="top") +
  ggtitle("NK Fitness Model")
```

```{r}
cor.test(x=data_nk_phase1$coding_sites, 
         y=data_nk_phase1$fitness,
         method="spearman",
         exact=FALSE)
```

## Publication Figures 

### Changing Environments Promote Modularity
```{r}
ggplot(data_gradient_phase0,
       aes(x=chg_rate_label, y=coding_sites)) +
  geom_boxplot() +
  # geom_jitter(alpha=0.1) +
  xlab("Environmental Change") +
  scale_y_continuous(name="Coding Sites\n(best genotype)",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  theme(legend.position="none") +
  ggsave("./imgs/coding-sites-v-chg-rate-gradient-phase0.pdf", width=10, height=7)
```

### Modularity Promotes Evolvability

```{r}
ggplot(data_gradient_phase1,
       aes(x=chg_rate_label, y=fitness)) +
  geom_boxplot() +
  xlab("Environmental Change") +
  scale_y_continuous(name="Fitness") +
  # ggtitle("Gradient Fitness Model (phase 2)") +
  theme(legend.position="none") + 
  ggsave("./imgs/fitness-v-chg-rate-gradient-phase1.pdf", width=10, height=7)
```

```{r}
ggplot(data_gradient_phase1,
       aes(x=coding_sites, y=fitness)) +
  geom_hex(alpha=0.95) +
  scale_fill_continuous(type = "viridis", 
                        trans="log",
                        name="Count",
                        breaks=c(1, 10, 100)) +
  scale_x_continuous(name="Coding Sites",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  ylab("Fitness") + 
  ggsave("./imgs/coding-sites-v-fitness-gradient-phase1-hex.pdf", width=10, height=7)
```

```{r}
ggplot(data_gradient_phase1,
       aes(x=coding_sites, y=fitness)) +
  geom_density_2d() +
  stat_density_2d(aes(fill = ..level..), 
                  geom = "polygon", 
                  colour="white") +
  geom_jitter(alpha=0.05) +
  scale_fill_continuous(type = "viridis", 
                        trans="log",
                        name="Count") +
  ylab("Fitness") +
  theme(legend.position = "none")
```

```{r}
ggplot(data_gradient_phase1,
       aes(x=coding_sites, y=fitness)) +
  geom_point(alpha=0.95) +
  scale_x_continuous(name="Coding Sites",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  ylab("Fitness")
```