---
title: "2020-07-07 - Does reduced mutation load drive compaction? - Experiment Analysis"
output:
  html_document:
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
---

## Analysis Dependencies

```{r, message=FALSE}
library(ggplot2)   # (Wickham, 2016)
library(tidyr)     # (Wickham and Henry, 2020)
library(dplyr)     # (Wickham et al., 2020)
library(reshape2)  # (Wickham, 2007)
library(cowplot)   # (Wilke, 2019)
library(patchwork) # (Pederson, 2020)
library(viridis)   # (Garnier, 2018)
library(hexbin)
```

We conducted these analyses using the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
data_path <- "./data/agg_data.csv"
agg_data <- read.csv(data_path, na.strings="NONE")

agg_data$BIT_FLIP_PROB <- as.factor(agg_data$BIT_FLIP_PROB)
agg_data$APPLY_BIT_MUTS_PER_GENE <- as.factor(agg_data$APPLY_BIT_MUTS_PER_GENE)
agg_data$DRIFT <- agg_data$TOURNAMENT_SIZE==1

chg_rate_label <- function(mag, interval, drift) {
  if (drift) { return("drift") }
  else if (interval == 0) { return("0") }
  else { return(paste(mag, interval, sep="/")) }
}

agg_data$chg_rate_label <- factor(mapply(chg_rate_label, 
                                            agg_data$CHANGE_MAGNITUDE,
                                            agg_data$CHANGE_FREQUENCY,
                                            agg_data$DRIFT),
                                     levels=c("drift", "0", "1/256", "1/128",
                                              "1/64", "1/32", "1/16", "1/8",
                                              "1/4", "1/2", "1/1", "2/1", 
                                              "4/1", "8/1", "16/1", "32/1",
                                              "64/1", "128/1", "256/1", 
                                              "512/1", "1024/1", "2048/1",
                                              "4096/1"))
theme_set(theme_cowplot())

data_nk_phase0 <- 
  filter(agg_data, GRADIENT_MODEL==0 & evo_phase == 0 & update==50000)
data_nk_phase1 <- 
  filter(agg_data, GRADIENT_MODEL==0 & evo_phase == 1 & update==60000)
data_gradient_phase0 <- 
  filter(agg_data, GRADIENT_MODEL==1 & evo_phase == 0 & update==50000)
data_gradient_phase1 <- 
  filter(agg_data, GRADIENT_MODEL==1 & evo_phase == 1 & update==60000)
```

## Evolution Experiment

### Fitness

#### Gradient fitness model

```{r}
ggplot(data_gradient_phase0, 
       aes(x=BIT_FLIP_PROB, y=fitness, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Fitness") +
  facet_wrap(~ chg_rate_label) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

#### NK fitness model

```{r}
ggplot(data_nk_phase0, 
       aes(x=BIT_FLIP_PROB, y=fitness, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Fitness") +
  facet_wrap(~ chg_rate_label) +
  ggtitle("NK Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

### Coding Sites

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase0, 
       aes(x=BIT_FLIP_PROB, y=coding_sites, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Coding Sites (in best organism)",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  facet_wrap(~ chg_rate_label) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90)) +
  ggsave("./imgs/gradient_phase0_coding_sites.pdf")
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase0, 
       aes(x=BIT_FLIP_PROB, y=coding_sites, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Coding Sites (in best organism)",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  facet_wrap(~ chg_rate_label) +
  ggtitle("NK Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90)) +
  ggsave("./imgs/nk_phase0_coding_sites.pdf")
```

### Genome Length

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase0, 
       aes(x=BIT_FLIP_PROB, y=genome_length, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Genome Length (in best organism)",
                     limits=c(0, 1025)) +
  facet_wrap(~ chg_rate_label) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase0, 
       aes(x=BIT_FLIP_PROB, y=genome_length, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Genome Length (in best organism)",
                     limits=c(0, 1025)) +
  facet_wrap(~ chg_rate_label) +
  ggtitle("NK Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

### Neutral Sites

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase0, 
       aes(x=BIT_FLIP_PROB, y=neutral_sites, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Neutral Sites (in best organism)") +
  facet_wrap(~ chg_rate_label) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase0, 
       aes(x=BIT_FLIP_PROB, y=neutral_sites, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Neutral Sites (in best organism)") +
  facet_wrap(~ chg_rate_label) +
  ggtitle("NK Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

## Transfer evolution experiment

### Fitness

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase1, 
       aes(x=BIT_FLIP_PROB, y=fitness, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Fitness") +
  facet_wrap(~ chg_rate_label) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase1, 
       aes(x=BIT_FLIP_PROB, y=fitness, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Fitness") +
  facet_wrap(~ chg_rate_label) +
  ggtitle("NK Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90))
```

### Fitness vs Coding Sites

#### Gradient Fitness Model

```{r}
ggplot(data_gradient_phase1, 
       aes(x=coding_sites, y=fitness, color=chg_rate_label)) +
  geom_jitter() +
  xlab("Coding Sites") +
  scale_y_continuous(name="Fitness") +
  facet_grid(APPLY_BIT_MUTS_PER_GENE ~ BIT_FLIP_PROB) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position = "bottom")
```

#### NK Fitness Model

```{r}
ggplot(data_nk_phase1, 
       aes(x=coding_sites, y=fitness, color=chg_rate_label)) +
  geom_jitter() +
  xlab("Coding Sites") +
  scale_y_continuous(name="Fitness") +
  facet_grid(APPLY_BIT_MUTS_PER_GENE ~ BIT_FLIP_PROB) +
  ggtitle("nk Fitness Model") +
  theme(legend.position = "bottom")
```

## Publication Figures

```{r}
ggplot(data_gradient_phase0, 
       aes(x=BIT_FLIP_PROB, y=coding_sites, color=APPLY_BIT_MUTS_PER_GENE)) +
  geom_boxplot() +
  xlab("Bit Flip Rate") +
  scale_y_continuous(name="Coding Sites (in best organism)",
                     limits=c(0, 130),
                     breaks=seq(0, 130, 20)) +
  facet_wrap(~ chg_rate_label) +
  ggtitle("Gradient Fitness Model") +
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=90)) +
  ggsave("./imgs/mutation_load_control_gradient.pdf")
```