---
title: "2020-06-14 - Is overlap adaptive at high mutation rates? Competition Experiments"
output:
  html_document:
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
---

## Analysis Dependencies

```{r, message=FALSE}
library(ggplot2)   # (Wickham, 2016)
library(tidyr)     # (Wickham and Henry, 2020)
library(dplyr)     # (Wickham et al., 2020)
library(reshape2)  # (Wickham, 2007)
library(cowplot)   # (Wilke, 2019)
library(patchwork) # (Pederson, 2020)
library(viridis)   # (Garnier, 2018)
library(hexbin)
```

We conducted these analyses using the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
data_path <- "./data/agg_data.csv"
agg_data <- read.csv(data_path, na.strings="NONE")

agg_data$BIT_FLIP_PROB <- as.factor(agg_data$BIT_FLIP_PROB)
agg_data$PHASE_2_BIT_FLIP_PROB <- as.factor(agg_data$PHASE_2_BIT_FLIP_PROB)
agg_data$DRIFT <- agg_data$TOURNAMENT_SIZE==1

chg_rate_label <- function(mag, interval, drift) {
  if (drift) { return("drift") }
  else if (interval == 0) { return("0") }
  else { return(paste(mag, interval, sep="/")) }
}

agg_data$chg_rate_label <- factor(mapply(chg_rate_label, 
                                            agg_data$CHANGE_MAGNITUDE,
                                            agg_data$CHANGE_FREQUENCY,
                                            agg_data$DRIFT),
                                     levels=c("drift", "0", "1/256", "1/128",
                                              "1/64", "1/32", "1/16", "1/8",
                                              "1/4", "1/2", "1/1", "2/1", 
                                              "4/1", "8/1", "16/1", "32/1",
                                              "64/1", "128/1", "256/1", 
                                              "512/1", "1024/1", "2048/1",
                                              "4096/1"))
agg_data$high_mut_arch_dominant <- 
  agg_data$low_mut_architecture_cnt <= agg_data$high_mut_architecture_cnt
agg_data$low_mut_arch_dominant <- 
  agg_data$low_mut_architecture_cnt >= agg_data$high_mut_architecture_cnt

theme_set(theme_cowplot())

data_nk <- 
  filter(agg_data, GRADIENT_MODEL==0)
data_gradient <- 
  filter(agg_data, GRADIENT_MODEL==1)

data_gradient_melted <- 
  melt(data_gradient,
     measure.vars=c("high_mut_arch_dominant",
                    "low_mut_arch_dominant"),
     variable.name="architecture_type",
     value.name="dominant")

data_nk_melted <- 
  melt(data_nk,
     measure.vars=c("high_mut_arch_dominant",
                    "low_mut_arch_dominant"),
     variable.name="architecture_type",
     value.name="dominant")

```

The graph is faceted by competition (columns) and bit substitution rate (rows).

```{r}
ggplot(data = data_gradient_melted, 
       aes(x = architecture_type, y = as.numeric(dominant), fill=architecture_type)) + 
  stat_summary(fun.y = sum, geom = "bar") +
  scale_x_discrete(breaks=c("high_mut_arch_dominant",
                            "low_mut_arch_dominant"),
                   labels=c("HM-Arch", "LM-Arch"),
                   name="Architecture Type") +
  facet_grid(BIT_FLIP_PROB~competition_category) +
  ggtitle("Gradient model") +
  theme(legend.position="none") +
  ggsave("./imgs/competition-gradient.pdf", width=10, height=10)
```

Counter-intuitive result with the gradient fitness model. 

Hypothesis: low mutation rate architecture better equiped to find high-fitness peaks early on at high mutation rate; high mutation rate architecture better equiped to quickly find local optima at low mutation rate (1 mutation can take bigger jumps in landscape).

```{r}
ggplot(data = data_nk_melted, 
       aes(x = architecture_type, y = as.numeric(dominant), fill=architecture_type)) + 
  stat_summary(fun.y = sum, geom = "bar") +
  scale_x_discrete(breaks=c("high_mut_arch_dominant",
                            "low_mut_arch_dominant"),
                   labels=c("HM-Arch", "LM-Arch"),
                   name="Architecture Type") +
  facet_grid(BIT_FLIP_PROB~competition_category) +
  ggtitle("NK model") +
  theme(legend.position="none") +
  ggsave("./imgs/competition-nk.pdf", width=10, height=10)
```